---
title: "p8105_hw6_jw4489"
output: github_document
date: "2023-12-02"
---

```{r, message = FALSE}
library(tidyverse)
library(purrr)
library(modelr)
library(dplyr)
library(broom)
library(ggplot2)
```

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE)
```

## Problem 1

```{r}
homicide = read.csv("hw6data/homicide-data.csv")
view(homicide)
```

##### Load and clean data

```{r}
df1 = homicide |>
  janitor::clean_names() |> 
  mutate(city_state = paste(city, state, sep = ", ")) |>
  mutate(solved = ifelse(disposition == "Closed by arrest", 1, 0)) |>
  filter(!(city %in% c("Dallas", "Phoenix", "Kansas City", "Tulsa"))) |>
  filter(victim_race == "White" | victim_race == "Black") |>
  mutate(victim_age = as.numeric(victim_age))

view(df1)  
```

##### Baltimore model

```{r}
baltimore_df = na.omit(df1) |>
   filter(city_state == "Baltimore, MD")
```

```{r}
baltimore_model = 
  glm(disposition == "Closed by arrest" ~ victim_age + victim_sex + victim_race, 
                      data = baltimore_df, family = binomial)
```

```{r}
baltimore_results = baltimore_model |>
  broom::tidy() |>
  filter(term == "victim_sexMale") |>
  mutate(estimate_or = exp(estimate), 
         CI_lower = exp(estimate - 1.96 * std.error),
         CI_upper = exp(estimate + 1.96 * std.error)) |>
  knitr::kable()      

baltimore_results
```

##### Each city

```{r}
city_results = df1 |>
  nest(data = -city_state) |>
  mutate(models = map(data, \(df) glm(solved ~ victim_age + victim_sex + victim_race, data = df, family = binomial())),
         results = map(models, broom::tidy)) |> 
  select(-models) |> 
  unnest(results) |>
  filter(term == "victim_sexMale") |>
  mutate(estimate_or = exp(estimate),
         CI_lower = exp(estimate - 1.96 * std.error),
         CI_upper = exp(estimate + 1.96 * std.error)) |>
  select(city_state, term, estimate_or, CI_lower, CI_upper)

city_results
```
  
##### Plots

```{r}
results_plot = city_results |>
  mutate(city_state = fct_reorder(city_state, estimate_or)) |> 
  ggplot(aes(x = city_state,y = estimate_or)) + 
  geom_point(size = 3) + 
  geom_errorbar(aes(ymin = CI_lower,ymax = CI_upper)) + 
  labs(x = "City", y = "Adjusted Odds Ratio (Male vs. Female)",
       title = "Adjusted Odds Ratios for Solving Homicides (Male vs. Female) by City") +
  theme(axis.text.x = element_text(size = 5, angle = 45, hjust = 1))

results_plot
```

## Problem 2

##### Load data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

##### Bootstrap function

```{r}
bootstrap_regression = function(data) {
  boot_sample = sample_n(data, nrow(data), replace = TRUE)
  model = lm(tmax ~ tmin + prcp, data = boot_sample)
  r_squared = broom::glance(model)$r.squared
  coefficients = broom::tidy(model) |>
    filter(term %in% c("tmin", "prcp")) |>
    summarise(log_beta_product = log(prod(estimate)))
  
  return(c(r_squared, coefficients))
}
```

##### Estimates of 5000 samples

```{r}
bootstrap_results <- data.frame(
  r_squared = numeric(5000),
  log_beta_product = numeric(5000)
)

set.seed(1) 
for (i in 1 : 5000) {
  bootstrap_results[i, ] <- bootstrap_regression(weather_df)
}
```

##### Visualizations

```{r}
ggplot(bootstrap_results, aes(x = r_squared)) +
  geom_histogram() +
  labs(x = "Bootstrap Estimates of r^2", y = "Frequency")

ggplot(bootstrap_results, aes(x = log_beta_product)) +
  geom_histogram() +
  labs(x = "Bootstrap Estimates of log(β1*β2)", y = "Frequency")
```

##### Confidence Interval

```{r}
quantile_0.025_1 = bootstrap_results |>
  pull(r_squared)|>
  quantile(0.025)

quantile_0.975_1 = bootstrap_results |>
  pull(r_squared)|>
  quantile(0.975)

quantile_0.025_1
quantile_0.975_1
```

According to the results above, the 95% confidence interval for r_squared is (`r quantile_0.025_1`, `r quantile_0.975_1`).

```{r}
quantile_0.025_2 = bootstrap_results |>
  pull(log_beta_product)|>
  na.omit()|>
  quantile(0.025)

quantile_0.975_2 = bootstrap_results |>
  pull(log_beta_product)|>
  na.omit()|>
  quantile(0.975)

quantile_0.025_2
quantile_0.975_2
```

According to the results above, the 95% confidence interval for og(β̂ 0∗β̂ 1) is (`r quantile_0.025_2`, `r quantile_0.975_2`).

## Problem 3

##### Load and clean the data

```{r}
birthweight = read.csv("hw6data/birthweight.csv")

view(birthweight)
```

```{r}
df3 = birthweight |>
  mutate(
    babysex = factor(babysex, levels = c(1, 2), labels = c("Male", "Female")),
    frace = factor(frace, levels = c(1, 2, 3, 4, 8, 9), labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
    mrace = factor(mrace, levels = c(1, 2, 3, 4, 8), labels = c("White", "Black", "Asian", "Puerto Rican", "Other"))) |>
  drop_na()

view(df3)
```

##### Regression model 1

For this part of the problem, I think the variables "babysex", "gaweeks", "malform", "wtgain",  "frace", and "mrace" can be the factors affecting the birthweight, so I use these variables to propose the regression model.

```{r}
model_1 = lm(bwt ~ babysex + gaweeks + malform + wtgain + frace + mrace, 
             data = df3) 
```

##### Plot

```{r}
plot_model_1 = df3 |>
  modelr::add_residuals(model_1) |> 
  modelr::add_predictions(model_1) |> 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  labs(x = "Fitted Values", y = "Residuals") +
  ggtitle("Residuals vs. Fitted Values")

plot_model_1
```

##### Compare with other two

Model 2: Using length at birth and gestational age as predictors (main effects only)

```{r}
model_2 = lm(bwt ~ blength + gaweeks, data = df3)
```

Model 3: Using head circumference, length, sex, and all interactions (including the three-way interaction)

```{r}
model_3 = lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex, data = df3)
```

```{r}
cv_df = crossv_mc(df3, 100) |>
  mutate(train = map(train, as_tibble), test = map(test, as_tibble)) |>
  mutate(model_1  = map(train, \(df) lm(bwt ~ babysex + gaweeks + malform + wtgain + frace + mrace, data = df3)),
        model_2  = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df3)),
        model_3 = map(train, \(df) lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex, data = df3))) |> 
  mutate(rmse_model_1 = map2_dbl(model_1, test, \(mod, df) rmse(model = mod, data = df)),
        rmse_model_2 = map2_dbl(model_2, test, \(mod, df) rmse(model = mod, data = df)),
        rmse_model_3 = map2_dbl(model_3, test, \(mod, df) rmse(model = mod, data = df)))

```

```{r}
plot_cv = cv_df |>
  select(starts_with("rmse")) |> 
  pivot_longer(everything(), names_to = "model", values_to = "rmse", names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin() +
  labs(x = "Model", 
       y = "Cross-Validated RMSE",
       title = "Comparison of three models")

plot_cv
```

